
# Install libraries 

```{r}
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(ggrepel) # For labels
library(viridis) # For colour palettes
```

# Graphs of LAI

## Upload LAI data from csv file

```{r}
LAI_data <- read.csv("LAI_2014.csv", check.names=FALSE, fileEncoding = 'UTF-8-BOM')
LAI_data
```
## Bar chart for 2014

### Reorder LAI data according to the index score

```{r}
LAI_data_reorder <- LAI_data # Replicate original data
LAI_data_reorder$country_name <- factor(LAI_data_reorder$country_name, # Factor levels in decreasing order
                  levels = LAI_data_reorder$country_name[order(LAI_data_reorder$D_LAI_2014, decreasing = TRUE)])
```

### Create graph of all European countries for 2014

```{r}
ggplot(LAI_data_reorder, aes(x= country_name, y = D_LAI_2014))+ 
  geom_col(stat = "identity", fill = "#D3D3D3")+
  ylim(0, 80)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Local Autonomy Index (LAI) for European countries (2014)", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```
## Bar chart for 2000

### Reorder LAI data according to the index score

```{r}
LAI_data_reorder_2000 <- LAI_data # Replicate original data
LAI_data_reorder_2000$country_name <- factor(LAI_data_reorder_2000$country_name, # Factor levels in decreasing order
                  levels = LAI_data_reorder_2000$country_name[order(LAI_data_reorder_2000$D_LAI_2000, decreasing = TRUE)])
```

### Create graph of all European countries for 2000

```{r}
ggplot(LAI_data_reorder_2000, aes(x= country_name, y = D_LAI_2000))+ 
  geom_col(stat = "identity", fill = "#D3D3D3")+
  ylim(0, 80)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Local Autonomy Index (LAI) for European countries (2000)", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```
## Bar chart for 1990

### Reorder LAI data according to the index score

```{r}
LAI_data_reorder_1990 <- LAI_data # Replicate original data
LAI_data_reorder_1990$country_name <- factor(LAI_data_reorder_1990$country_name, # Factor levels in decreasing order
                  levels = LAI_data_reorder_1990$country_name[order(LAI_data_reorder_1990$D_LAI_1990, decreasing = TRUE)])
```

### Create graph of all European countries for 1990

```{r}
ggplot(LAI_data_reorder_1990, aes(x= country_name, y = D_LAI_1990))+ 
  geom_col(stat = "identity", fill = "#D3D3D3")+
  ylim(0, 80)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Local Autonomy Index (LAI) for European countries (1990)", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```

### Export as tiff image

# Bar chart for change in LAI

```{r}
LAI_data_reorder_change <- LAI_data # Replicate original data
LAI_data_reorder_change$country_name <- factor(LAI_data_reorder_change$country_name, # Factor levels in decreasing order
                  levels = LAI_data_reorder_change$country_name[order(LAI_data_reorder_change$D_LAI_changes_1990_2014, decreasing = TRUE)])
```

```{r}
ggplot(LAI_data_reorder_change, aes(x= country_name, y = D_LAI_changes_1990_2014))+ 
  geom_col(stat = "identity", fill = "#D3D3D3")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Change in Local Autonomy Index (LAI) for European countries (1990-2014)", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```
## Bar chart for 2010

### Reorder LAI data according to the index score

```{r}
LAI_data_reorder_2010 <- LAI_data # Replicate original data
LAI_data_reorder_2010$country_name <- factor(LAI_data_reorder_2010$country_name, # Factor levels in decreasing order
                  levels = LAI_data_reorder_2010$country_name[order(LAI_data_reorder_2010$D_LAI_2010, decreasing = TRUE)])
```

### Create graph of all European countries for 2010

```{r}
ggplot(LAI_data_reorder_2010, aes(x= country_name, y = D_LAI_2010))+ 
  geom_col(stat = "identity", fill = "#D3D3D3")+
  ylim(0, 80)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title="Local Autonomy Index (LAI) for European countries (2010)", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```

# Grouped barchart for all years

```{r}
 # List of column names
colnames(LAI_data)

# Select specific countries
# Rename years
LAI_data_select <- LAI_data %>%
  select("country_name", "D_LAI_1990", "D_LAI_2000", "D_LAI_2010", "D_LAI_2014") %>%
  filter(country_name %in% c("Germany", "France", "United Kingdom", "Italy", "Spain", "Belgium", "Netherlands", "Portugal", "Ireland")) %>% # Select specific countries
  rename("1990" = D_LAI_1990, "2000" = D_LAI_2000, "2010" = D_LAI_2010, "2014" = D_LAI_2014) # Rename years
```

```{r}
# Make data long
LAI_data_select_long <-gather(LAI_data_select ,key="year",value="LAI", c("1990", "2000", "2010", "2014"))
print(LAI_data_select_long)
```

### In black and white

```{r}
ggplot(LAI_data_select_long, aes(y=LAI, x=country_name)) + 
    geom_bar(position="dodge", stat="identity", aes(fill = year))+
  scale_fill_manual(values = c("1990" = "#cccccc",
                               "2000" = "#969696",
                               "2010" = "#636363",
                               "2014" = "#252525"))+
  ylim(0, 80)+
  theme_minimal()+
  theme(legend.title = element_blank())+
  labs(title="Local Autonomy Index (LAI) for selected European countries", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```

## In colour

```{r}
ggplot(LAI_data_select_long, aes(y=LAI, x=country_name)) + 
    geom_bar(position="dodge", stat="identity", aes(fill = year))+
  scale_fill_manual(values = c("1990" = "#ffffcc",
                               "2000" = "#a1dab4",
                               "2010" = "#41b6c4",
                               "2014" = "#225ea8"))+
  ylim(0, 80)+
  theme_minimal()+
  theme(legend.title = element_blank())+
  labs(title="Local Autonomy Index (LAI) for selected European countries", y="LAI", x = " ", caption = "Index based on Ladner et al. (2015)")
```
